---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# learntidymodels

<!-- badges: start -->
[![R build status](https://github.com/tidymodels/learntidymodels/workflows/R-CMD-check/badge.svg)](https://github.com/tidymodels/learntidymodels/actions)
<!-- badges: end -->

Learn [tidymodels](https://www.tidymodels.org/) in the browser or locally in 
your RStudio IDE with interactive [learnr](https://rstudio.github.io/learnr/) primers!

## Installation

For now, you can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("tidymodels/learntidymodels")
```
## How to run the tutorials

To run the tutorials, you should also install the `learnr` package. Once both `learntidymodels` and `learnr` packages are installed, you can easily start any tutorial with:

``` r
install.packages("learnr")
learnr::run_tutorial("tutorial_of_choice", package = "learntidymodels")
```

For example:

``` r
learnr::run_tutorial("PCA_recipes", package = "learntidymodels")
```

## List of available tutorials

```{r, eval=TRUE, echo=FALSE, message=FALSE, error=FALSE, warning=FALSE}
knitr::kable(readr::read_tsv("tutorial_list.tsv"))
```


## Code of Conduct
  
Please note that this project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
  
