---
title: "Dimension Reduction with recipes"
output: 
  learnr::tutorial:
    progressive: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE, message=FALSE}
library(learnr)
library(tidymodels)
library(tidyverse)
library(embed)
set.seed(9876)


knitr::opts_chunk$set(echo = FALSE)

zoo_names <- scan(text="animal_name
hair
feathers
eggs
milk
airborne
aquatic
predator
toothed
backbone
breathes
venomous
fins
legs
tail
domestic
catsize
class", what="character")
anim_types <- tribble(~class, ~type,
                      1, "mammal",
                      2, "bird",
                      3, "reptile",
                      4, "fish",
                      5, "amphibian",
                      6, "insect",
                      7, "other_arthropods")

zoo <- 
  read_csv("http://archive.ics.uci.edu/ml/machine-learning-databases/zoo/zoo.data", 
           col_names = zoo_names) %>%
  left_join(anim_types) %>%
  select(-class)
# exercise.lines = 5


pca_rec <- recipe(data = zoo, formula = ~.) %>%
  update_role(animal_name, type, new_role = "id") %>%
  step_normalize(all_predictors()) %>%
  step_pca(all_predictors())
pca_prep <- prep(pca_rec)
pca_juice <- juice(pca_prep)
```


## Welcome

Dimension reduction is a regularly used unsupervised method in exploratory data analysis and predictive models.

This tutorial will teach you how to apply these methods using [recipes](https://recipes.tidymodels.org/) package, which is a part of the [tidymodels](https://www.tidymodels.org) metapackge.

If you are new to [tidymodels](https://www.tidymodels.org), you can learn what you need with the five [Get Started articles on tidymodels.org](https://www.tidymodels.org/start/). Second article, [Preprocessing your data with recipes](https://www.tidymodels.org/start/recipes/), shows many great things you can do with [recipes](https://recipes.tidymodels.org/).

The tutorial focuses on two algorithms:

1. Linear dimensionality reduction with [Principal component analysis (PCA)](https://en.wikipedia.org/wiki/Principal_component_analysis)   
2. Non-linear dimensionality reduction with [UMAP](https://umap-learn.readthedocs.io/en/latest/how_umap_works.html)

Using [recipes](https://recipes.tidymodels.org/) we will apply these methods to explore our data. However, these methods can also be used for [feature extraction](https://en.wikipedia.org/wiki/Feature_extraction) prior to modeling. 

Let's get started!

## Enter the `zoo`

We will use the `zoo` dataset to explore these methods. `zoo` contains observations collected on `r nrow(zoo)` zoo animals.

To see the first ten rows of the dataset click "Run Code".

```{r intro-zoo, exercise=TRUE}
zoo
```

You can use the black triangle that appears at the top right of the table to scroll through all of the columns in `zoo`.

Alternatively, use `glimpse()` to see columns in a more compact way:

```{r glimpse-zoo, exercise=TRUE}

```


```{r glimpse-zoo-solution}
glimpse(zoo)
```


We can see that `zoo` includes variables such as:

1. `animal_name`, name of the animal observed
2. `type`, type of the animal   
3. And `r ncol(zoo)-2` other characteristics such as: `hair`, if the animal has hair (coded in `1` for yes and `0` for no), or `legs`, how many legs does the animal have, etc.

### Explore the `zoo`

Let's see how many animals were observed from each type.
Click "Run Code" to get the answer.

```{r type-bar, exercise=TRUE}
zoo %>%
  ggplot(aes(type)) +
  geom_bar()
```

We can see that there are many mammals and birds, but fewer reptiles and amphibians in `zoo`.

We can also look at the distribution of animals that lay `eggs` across animal types. Click `Run Code` to generate the plot. 

```{r type-eggs-bar, exercise=TRUE}
zoo %>%
  mutate(eggs=recode(eggs, `1`="Lays Eggs", `0`="Doesn't lay eggs")) %>%
  ggplot(aes(type, fill=eggs)) +
  geom_bar()
```

We can see that majority of the mammals don't produce eggs. 
Can you find the one mammal that does?

```{r find-eggs, exercise=TRUE}
# Remember dplyr package? 
# You can use filter() to choose rows/cases where conditions are true.

```

```{r find-eggs-hint, eval=FALSE}
___ %>%
  filter(___ == ___)
```

```{r find-eggs-solution, eval=FALSE}
zoo %>%
  filter(type == "mammal",
         eggs == 1)
```


<!-- Let's see how these animal features correlate with each other. -->
<!-- Run the code to plot the correlation matrix using the [corrr](https://cran.r-project.org/web/packages/corrr/vignettes/using-corrr.html) package. -->

<!-- ```{r} -->
<!-- library(corrr) -->
<!-- zoo %>% -->
<!--   select(-animal_name, -type) %>% -->
<!--   correlate() %>% -->
<!--   rearrange() %>% -->
<!--   rplot(shape = 15, colours = c("darkorange", "white", "darkcyan"), print_cor=TRUE)  -->
<!-- ``` -->

<!-- We can see that producing eggs or milk have a very strong negative correlation. (The only reason it isn't equal to `-1` is the odd ball platypus.) -->


## Dimension Reduction with **recipes**

We would like to explore the relationship of these animal features in the data and how similar animals are to each other. Potentially, we would like to see if same `type` of animals will _cluster closely_ upon dimension reduction.

### Principal Component Analysis

Let's start by implementing principal component analysis (PCA) using [recipes](https://recipes.tidymodels.org/) from [tidymodels](https://www.tidymodels.org).

First, we will create our **recipe**.

Let's go through the steps we took to create this **recipe**

1. Define a dataset and _formula_ using `recipe()`      
  We provide our data `zoo` and our formula `~.`. Notice that we do not define an outcome and use `.` in our formula to select all variables for the next steps of the analysis.    
  
2. Change the _role_ of identifier variables with `update_role()`     
  While we already told recipe to include all variables, we want to exclude identifier columns `animal_name` or `type` from our analysis. On the other hand we would like to keep these columns for the next steps when we are plotting our results. By using `update_role()` we achieve this without completely dropping these variables for the next steps.   
  
3. Normalize variables with `step_normalize`   
  With this step we tell recipe to normalize numeric data to have a standard deviation of one and a mean of zero. We also use the helper function `all_predictors()` to select all the variables we would like to normalize    
  
4. Finally, we tell the recipe to convert all variables (except `animal_name` and `type`) into principal components with `step_pca()`.


Click "Run Code" to create a recipe object and print it.

```{r pca-rec, exercise=TRUE, exercise.lines=6}
pca_rec <- recipe(data = zoo, formula = ~.) %>%
  update_role(animal_name, type, new_role = "id") %>%
  step_normalize(all_predictors()) %>%
  step_pca(all_predictors())

pca_rec
```

Not the expected output?   

Because so far we only _defined_ our recipe, but did not run or implement it. To get the results from our PCA, we need evaluate our **recipe** using `prep()`.
  
```{r pca-prep, exercise=TRUE}
pca_prep <- prep(pca_rec)
pca_prep
```

Can you see the difference between the outputs? After `prep()` we can see that centering for, and PCA extraction with all columns of interest has been completed or "[trained]".

We can also look at the steps this recipe contains with `tidy()`:

```{r tidy-prep, exercise=TRUE}
tidy(pca_prep)
```

We can see that two `type` of steps are contained in this prepped recipe: (1) `normalize` and (2) `pca`.

Now, in order to extract the values for plotting, we need to use `juice()`. This way we can extract the step `pca`.

Click "Run Code" to plot the first and second principal components.

```{r juice-plot, exercise=TRUE}
pca_juice <- juice(pca_prep)
pca_juice

pca_juice %>%
  ggplot(aes(PC1, PC2, label=animal_name)) +
  geom_point(aes(color = type), alpha = 0.7, size = 2)+
  geom_text(check_overlap = TRUE, hjust = "inward") +
  labs(color = NULL)
```


### UMAP

Now do you think you can take the same steps to make a plot with UMAP algorithm? You can follow the pretty much the same steps.

Some hints:

1. Change `step_pca()` to `step_umap`
2. When plotting, swap `PC1, PC2` with `umap_1, umap_2`

Now, give it a go!


```{r umap-try, exercise=TRUE, exercise.lines=12}

```


```{r umap-try-solution, eval=FALSE}
umap_rec <- recipe(~., data = zoo) %>%
  update_role(animal_name, type, new_role = "id") %>%
  step_normalize(all_predictors()) %>%
  step_umap(all_predictors())

umap_prep <- prep(umap_rec)

juice(umap_prep) %>%
  ggplot(aes(umap_1, umap_2, label=animal_name)) +
  geom_point(aes(color = type), alpha = 0.7, size = 2)+
  geom_text(check_overlap = TRUE, hjust = "inward") +
  labs(color = NULL)
```

<!-- ### Quiz -->

<!-- *You can include any number of single or multiple choice questions as a quiz. Use the `question` function to define a question and the `quiz` function for grouping multiple questions together.* -->

<!-- Some questions to verify that you understand the purposes of various base and recommended R packages: -->

<!-- ```{r quiz} -->
<!-- quiz( -->
<!--   question("Which package contains functions for installing other R packages?", -->
<!--     answer("base"), -->
<!--     answer("tools"), -->
<!--     answer("utils", correct = TRUE), -->
<!--     answer("codetools") -->
<!--   ), -->
<!--   question("Which of the R packages listed below are used to create plots?", -->
<!--     answer("lattice", correct = TRUE), -->
<!--     answer("tools"), -->
<!--     answer("stats"), -->
<!--     answer("grid", correct = TRUE) -->
<!--   ) -->
<!-- ) -->
<!-- ``` -->

